Parsing Postfix log files

### 2008-11-23 doc/logparser-background.tex
Started updating blurbs etc. in the start of the art review. Dealt with a few XXX items.

This section provides background information helpful in understanding the
remainder of the thesis.  It begins with a discussion of the motivation
underlying the project, followed by some technical information: the use of
a database as an \gls{API}\@; a brief introduction to \gls{SMTP}\@; a
longer introduction to Postfix, concentrating on anti-spam restrictions and
policy servers.  The assumptions made in designing and implementing the
parser are explained, as are the conventions used in this thesis.

@@

\subsection{Conventions used}

XXX MAYBE REWRITE SO THAT THIS IS NOT TRUE\@?  ARE THERE ANY OTHER
CONVENTIONS\@?

The words \textit{connection\/} and \textit{mail\/} are often used
interchangeably; in general the word used was chosen based
on the context it appears in.

@@

This section has provided background information on several topics,
starting with the motivation behind the project, continuing with an
explanation of the use of a database as an \gls{API} and an introduction to
\gls{SMTP}.  The focus then switched to the \gls{MTA} in use, Postfix, and
its restrictions and policy servers, finishing with the conventions used in
this thesis.

>>>>

This section provides background information helpful in understanding the
remainder of the thesis.  It begins with a discussion of the motivation
underlying the project, followed by some technical information: the use of
a database as an \gls{API}\@; a brief introduction to \gls{SMTP}\@.  It
finishes with a longer introduction to Postfix, concentrating on anti-spam
restrictions and policy servers.

@@

@@

This section has provided background information on several topics,
starting with the motivation behind the project, continuing with an
explanation of the use of a database as an \gls{API} and an introduction to
\gls{SMTP}.  The focus then switched to Postfix, the \gls{MTA} in use, and
its restrictions and policy servers.

### 2008-11-23 doc/logparser-other-parsers-review.tex
Started updating blurbs etc. in the start of the art review. Dealt with a few XXX items.

\noindent{}\url{http://jimsun.linxnet.com/postfix_contrib.html} \newline{}
(Last checked 2008/04/09.)

@@

    Sawmill is a Postfix log analyzer (it also support [sic] 686 other log
    formats).  It can process log files in Postfix format, and generate
    dynamic statistics from them, analyzing and reporting events.  Sawmill
    can parse Postfix logs, import them into a SQL database (or its own
    built-in database), aggregate them, and generate dynamically filtered
    reports, all through a web interface.  Sawmill can perform Postfix
    analysis on any platform, including Window, Linux, FreeBSD, OpenBSD,
    Mac OS, Solaris, other UNIX, and more.

@@

\noindent{}\url{http://www.thesawmill.co.uk/formats/postfix.html}
\newline{} \url{http://www.thesawmill.co.uk/formats/postfix_ii.html}
\newline{} \url{http://www.thesawmill.co.uk/formats/beta_postfix.html}
\newline{} (Last checked 2007/12/04.)

XXX CHECK ALL THIS AGAIN\@; GIVE EXAMPLE PRICES\@.

Sawmill is a general purpose commercial product that parses 687 log file
formats (correct as of 2007/12/04) and produces reports.  Its data
extraction facilities are too limited to save sufficient data for the
purposes of this project,\footnote{The data extracted by Sawmill is
described later in this review, and the data extracted by \gls{PLP} is
described in \sectionref{connections table} and \sectionref{results
table}.} though it does extract three different sets of data for Postfix
(one is beta as of 2007/12/04), but they do not appear to be interlinked.
The documentation does not suggest that any attempt is made to correlate
log lines or deal with the difficulties documented in
\sectionref{complications}.  The source code is available in an obfuscated
form only (presumably for a fee), and the product is quite expensive,
requiring a license per report to be generated; in contrast \parsername{}
is free and the code is freely available.  The web interface allows
creation of dynamic reports based on any field, but because of the
associated cost the author has not experimented with it.  Presumably, as
the company charges per report the user wishes to generate, the data store
(if there is one) is deliberately inaccessible and undocumented to prevent
the user bypassing the program and generating their own reports.

XXX THE BLURB MENTIONS \gls{SQL} DB\@; CAN THAT BE ACCESSED SEPARATELY, IS
THE FORMAT DOCUMENTED, ETC\@?

\url{http://www.thesawmill.co.uk/formats/postfix.html} \newline{} Fields
extracted: from, to, server, UID, relay, status, number of recipients,
origin hostname, origin \gls{IP}, and virus.  The fields \texttt{server},
\texttt{uid} and \texttt{virus} are not explained in the documentation:
\texttt{server} is probably the server the mail is delivered to, and
\texttt{uid} might be the uid of the user submitting mail locally.  Postfix
does not perform any form of virus checking (though it has many options for
cooperating with an external virus scanner), so the \texttt{virus} field is
a mystery.

\url{http://www.thesawmill.co.uk/formats/postfix_ii.html} \newline{} Fields
extracted: from, to, RBL, client hostname, and client \gls{IP}\@.

\url{http://www.thesawmill.co.uk/formats/beta_postfix.html} \newline{}
Fields extracted: from, to, client hostname, client \gls{IP}, relay
hostname, relay \gls{IP}, status, response code, RBL, and message id.

Even if the three data sets were linked together Sawmill would extract less
data than \parsername{}, and it does not appear to extract data about
rejections except when the rejection is caused by a \gls{DNSBL} check.

@@

@@

    Splunk is an IT Search engine. It is software that indexes any format
    of IT data from any source in real time, including logs,
    configurations, scripts, code, messages, traps, alerts, activity
    reports, stack traces and metrics from all of your applications,
    servers and devices. Splunk lets you search, navigate, alert and report
    on all your IT data in real time using an AJAX web interface. You can
    also share knowledge and Splunk solutions with other members of the
    Splunk community via SplunkBase.

@@

\noindent{}\url{http://www.splunk.com/} \newline{}
(Last checked 2008/04/29.)

>>>>

\noindent{}\url{http://jimsun.linxnet.com/postfix_contrib.html} \newline{}
(Last checked 2008/11/23.)

@@

    Sawmill is a Postfix log analyzer (it also support 818 other log
    formats). It can process log files in Postfix format, and generate
    dynamic statistics from them, analyzing and reporting events. Sawmill
    can parse Postfix logs, import them into a SQL database (or its own
    built-in database), aggregate them, and generate dynamically filtered
    reports, all through a web interface. Sawmill can perform Postfix
    analysis on any platform, including Window, Linux, FreeBSD, OpenBSD,
    Mac OS, Solaris, other UNIX, and more.

@@

\noindent{}\url{http://www.thesawmill.co.uk/formats/postfix.html}
\newline{} \url{http://www.thesawmill.co.uk/formats/postfix_ii.html}
\newline{} \url{http://www.thesawmill.co.uk/formats/beta_postfix.html}
\newline{} (Last checked 2008/11/23.)

Sawmill is a general purpose commercial product that parses 818 log file
formats (as of 2008/11/23) and produces reports from the extracted data.
Its data extraction facilities (described later) are too limited to save
sufficient data for the purposes of this project; it extracts three
different sets of data from Postfix log files, but they are not interlinked
in any way.  The documentation does not suggest that any attempt is made to
correlate log lines or deal with the difficulties documented in
\sectionref{complications}.

Sawmill has three different Postfix log file parsers:

    \item \url{http://www.thesawmill.co.uk/formats/postfix.html} \newline{}
        Fields extracted: from, to, server, UID, relay, status, number of
        recipients, origin hostname, origin \gls{IP}, and virus.  It also
        counts the number of and total size of all mails delivered.  The
        fields \texttt{server}, \texttt{uid}, \texttt{relay}, and
        \texttt{virus} are not explained in the documentation:
        \texttt{server} is probably the server the mail is delivered to,
        and \texttt{uid} might be the uid of the user submitting mail
        locally.  Postfix does not perform any form of virus checking
        (though it has many options for cooperating with an external virus
        scanner), so the \texttt{virus} field is a mystery.

    \item \url{http://www.thesawmill.co.uk/formats/postfix_ii.html}
        \newline{} Fields extracted: from, to, RBL list, client hostname,
        and client \gls{IP}\@.  It also counts the number of and total size
        of all mails delivered.

    \item \url{http://www.thesawmill.co.uk/formats/beta_postfix.html}
        \newline{} Fields extracted: from, to, client hostname, client
        \gls{IP}, relay hostname, relay \gls{IP}, status, response code,
        RBL list, and message id.  It also counts the number and size of
        all mails delivered, processed, blocked, expired, and bounced.

Even if the three data sets were linked together Sawmill would extract less
data than \parsername{}, and it does not appear to extract data about
rejections except when the rejection is caused by a \gls{DNSBL} check
(\textit{RBL list\/} in the list of fields).

@@

XXX THE BLURB MENTIONS \gls{SQL} DB\@; CAN THAT BE ACCESSED SEPARATELY, IS
THE FORMAT DOCUMENTED, ETC\@?  XXX EXPERIMENT WITH THE REPORTS IN THE DEMO
VERSION\@.  XXX EVERYTHING FROM HERE DOWN NEEDS TO BE CHECKED AND/OR
REWRITTEN\@.

The web interface allows creation of dynamic reports based on any field,
but because of the associated cost the author has not experimented with it.
As the company charges per report the user wishes to generate, the data
store (if there is one) is deliberately inaccessible and undocumented to
prevent the user bypassing the program and generating their own reports.

The source code is available only in an obfuscated form (presumably for a
fee), and the product is quite expensive, requiring a \euros{100} + VAT
license per report to be generated; in contrast \parsername{} is free and
the code is freely available.

@@

    Splunk is IT Search.

    Search and navigate IT data from applications, servers and network
    devices in real-time. Logs, configurations, messages, traps and alerts,
    scripts, code, metrics and more. If a machine can generate it ---
    Splunk can eat it. It's easy to download and use and it's very
    powerful.

@@

\noindent{}\url{http://www.splunk.com/} \newline{}
(Last checked 2008/11/23.)
